---
title: "Case Study 'Tag der Befreiung' in GermaParl"
editor: source
author: Philipp Heinrich
date: May 18, 2025
format:
  html:
    toc: true
    toc-float: true
    collapsed: false
    smooth-scroll: false
    code-fold: true
    number-sections: true
    df-print: paged
---

## notation
- "Befreiung" refers to a word or lemma
- **Befreiung** is a discourseme

## setup

- a couple of helper functions are defined in cwb-cads/R-utils
```{r, message = FALSE}
source("R-utils/collocation.R")
source("R-utils/concordance.R")
source("R-utils/association.R")
source("R-utils/api.R")
library(corpora)
```

### API

- beta server
```{r}
cads_connect(user="admin", password="tag-der-befreiung")
```

### corpus

```{r}
corpora <- cads_list_corpora()
corpora
```

- we use GermaParl v2.2.0
- data set covers all records of the German parliament from 1949-2024 (legislative periods (LPs) 1 through 20)
- NB we use original GERMAPARL220 here, not GERMAPARL220WEB
- final data set comprises 288,054,199 tokens in 4524 plenary protocols and 1,018,495 speeches

```{r}
cads_activate_corpus("GERMAPARL220")
```

### meta data

```{r}
meta <- cads_corpus_metadata()
meta
```

#### parliamentary groups

```{r}
cads_metadata_frequencies("speaker", "parlgroup") |>
  extract2("frequencies") ->
  meta.groups
```

- note that far too most entries are n/a, so we have a very incomplete classification if we rely on this level

```{r}
meta.groups |> 
  ggplot(aes(y = bin_unicode, x = nr_spans)) +
  geom_col()
```

#### parties

```{r}
cads_metadata_frequencies("speaker", "party") |>
  extract2("frequencies") -> 
  meta.all.parties
```

```{r}
meta.all.parties |> 
  ggplot(aes(y = bin_unicode, x = nr_spans)) +
  geom_col()
```

- reduce to major relevant parties for visualisations, recoding everything else als _andere_

```{r}
map.parties <- function (x, ordered=FALSE) {
  lvl <- qw("AfD,CSU,CDU,FDP,SPD,GRUENE,PDS,DIE LINKE,andere", sep=",")
  x <- ifelse(x %in% lvl, x, "andere")
  if (ordered) factor(x, levels=lvl) else x
}
meta.all.parties |> 
  mutate(party=map.parties(bin_unicode)) |>
  group_by(party) |>
  summarize(nr_spans=sum(nr_spans), nr_tokens=sum(nr_tokens)) ->
  meta.parties
```

```{r}
meta.parties |> 
  ggplot(aes(y=map.parties(party, ordered=TRUE), x=nr_tokens/1e6)) +
  geom_col() + ylab("Partei") + xlab("M Token")
```


#### years

```{r}
to.date <- function (x) as.Date(str_c(x, "-01-01"))
cads_metadata_frequencies("protocol", "year") |>
  extract2("frequencies") |>
  mutate(year = to.date(bin_unicode)) |>
  tibble() -> meta.years
```

```{r}
meta.years |> 
  ggplot(aes(x = year, y = nr_tokens / 1e6)) +
  geom_col() + ylab("M Token") + xlab("Jahr")
```



### subcorpora

```{r}
subcorpora <- cads_list_subcorpora()
subcorpora
```


#### create subcorpora

- I already created subcorpora for parliamentary groups (and years)

```{r}
groups <- c("CDU/CSU", "GRUENE", "SPD", "AfD", "DIE LINKE", "FDP")
subcorpora <- groups
```

- you can create further subcorpora via PUT request on subcorpus endpoint
- NB this endpoint has been implemented with datetime variables in mind but should work fine for unicode meta data
- NB you can create complex subcorpora (e.g. SPD in LP19) iteratively (i.e. via a subcorpus on a subcorpus)

**TODO** API for subcorpus info and working with subcorpora

- subcorpus collections (e.g. all parties) can be created via following endpoint
- **TODO** create API for this

```{r}
# collection <- str_interp("${url_api}/corpus/${corpus.id}/subcorpus-collection/") |>
#   request() |> 
#   req_auth_bearer_token(access.token) |>
#   req_body_json(list(name = 'parliamentary groups', level = 'speaker', key = 'parlgroup')) |> 
#   req_method('put') |> 
#   req_perform() |> 
#   magrittr::extract2("body") |> rawToChar() |> fromJSON()
```


## corpus queries

### query & concordance

- various queries have already be created

```{r}
cads_list_queries()
```

- since it's more or less impossible to find a specific query in this table (because the `cqp_query` is often cryptic with ad-hoc wordlists and there is no meaningful label), we have to run a new query, adding yet another entry to the list
- we first look at "Tag der Befreiung" and related terms

```{r}
query <- cads_run_simple_query(list("Tag die Befreiung", "8. Mai", "8. Mai 1945", "8. Mai die Jahr 1945"), p.att="lemma_smor")
query.id <- query$id
```

- query object returned contains various information, most importantly the query ID

```{r}
query
```

- retrieve concordance for this query (first page of 25 items) and use convenience function `conc.kwic()` for display

```{r}
concordance <- cads_query_concordance(query.id, page.size=25, window=5)
concordance |> conc.kwic()
```

- there's a substantial amount of FPs since May 8th can indeed refer to other things (d'uh), we will ignore this issue here

- we can also obtain a frequency breakdown of the query matches

```{r}
breakdown <- cads_query_breakdown(query.id, p.att="lemma_smor")
breakdown$items |> arrange(desc(freq))
```

### Tag der Befreiung

- we want to plot distribution of the expression "Tag der Befreiung" ➞ need to obtain metadata from concordance and aggregate frequencies

```{r}
tdb.query <- cads_run_simple_query(list("Tag die Befreiung"), p.att="lemma_smor")
tdb.query.id <- tdb.query$id
```

- pull entire concordance so we can extract metadata

```{r}
tdb.conc <- cads_query_concordance(tdb.query.id, secondary.att="lemma_smor", page.size=1e6)
```

```{r}
tdb.conc$lines$structural |> 
  select(protocol_date, protocol_year, speaker_name, speaker_party) |>
  mutate(year=to.date(protocol_year), party=map.parties(speaker_party)) ->
  tdb.meta
```

- compute relative frequencies for parties and years and display as barplots

```{r}
tdb.meta |> 
  summarise(f=n(), .by="year") |>
  right_join(meta.years, by="year") |>
  mutate(pmw=1e6 * coalesce(f, 0) / nr_tokens) |>
  ggplot(aes(y=pmw, x=year)) +
  geom_col() + ylab("PMW") + xlab("Jahr") + ylim(c(0, 5))
```

```{r}
tdb.meta |> 
  summarise(f=n(), .by="party") |>
  right_join(meta.parties, by="party") |>
  mutate(pmw=1e6 * coalesce(f, 0) / nr_tokens) |>
  ggplot(aes(x=pmw, y=party)) +
  geom_col() + ylab("Partei") + xlab("PMW") + xlim(c(0, 5))
```

## try to work with discoursemes in a constellation

```{r}
constellations <- cads_list_constellations()
constellations
```

- work with the latest constellation (presumably from last version of PH's script)

```{r}
constellation.id <- 19
constellations |> filter(id == constellation.id) |> pull("discoursemes") |> extract2(1) -> discoursemes19
discoursemes19
```

- now we can display the discoursemes

```{r}
for (i in seq_len(nrow(discoursemes19))) {
  discoursemes19$template[[i]] |> mutate(discourseme=discoursemes19$name[i]) |> print()
}
```

```{r}
str_interp("mmda/constellation/${constellation.id}/description/") |>
  cads_mk_request() |>
  cads_perform_request() ->
  descriptions19

print(descriptions19)
print(descriptions19$discourseme_descriptions[[1]])
```


## focus discourseme

### creation
- we create focus discourseme using lemma "Befreiung"
```{r}
focus.discourseme <- str_interp("${url_api}/mmda/discourseme/") |> 
  request() |> 
  req_auth_bearer_token(access.token) |> 
  req_body_json(list(name = "Befreiung")) |> 
  req_method("POST") |>
  req_perform() |> 
  magrittr::extract2("body") |>
  rawToChar() |> fromJSON()

focus.discourseme.id <- focus.discourseme |> magrittr::extract2("id")
```

- and a corresponding discourseme description
```{r}
focus.discourseme.description <- str_interp("${url_api}/mmda/discourseme/${focus.discourseme.id}/description/") |> 
  request() |> 
  req_auth_bearer_token(access.token) |> 
  req_body_json(
    list(corpus_id = corpus.id, 
         items = list(
           list(p = "lemma", surface = "Befreiung")
         ))
    ) |>
  req_perform() |> 
  magrittr::extract2("body") |>
  rawToChar() |> fromJSON()

focus.discourseme.description.id <- focus.discourseme.description |> magrittr::extract2("id")
```

### breakdown
```{r}
breakdown <- str_interp("${url_api}/mmda/discourseme/${focus.discourseme.id}/description/${focus.discourseme.description.id}/breakdown") |> 
  request() |> 
  req_auth_bearer_token(access.token) |>
  req_url_query(p = "lemma") |> 
  req_perform() |> 
  magrittr::extract2("body") |> rawToChar() |> fromJSON() |> 
  magrittr::extract2("items") |> 
  tibble()

breakdown |> rename(lemma = item) |> 
  arrange(desc(ipm)) |> 
  mutate(ipm = round(ipm, 2),
         ratio = round(freq / sum(freq), 2)) |> 
  arrange(desc(ipm)) |> 
  select(lemma, freq, ipm, ratio)
```

### supervised extension

- TODO doesn't work on the beta server ("Truncated or oversized response headers received from daemon process 'mmda-beta': /data/Philipp/cwb-cads-beta/cwb-cads.wsgi")

### frequencies across groups, years, and periods

```{r}
meta.dist <- str_interp("${url_api}/mmda/discourseme/${focus.discourseme.id}/description/${focus.discourseme.description.id}/meta") |> 
  request() |> 
  req_url_query(level = "speaker", key = "parlgroup", page_size = 1000) |> 
  req_auth_bearer_token(access.token) |> 
  req_perform() |> 
  magrittr::extract2("body") |>
  rawToChar() |> fromJSON()
```

```{r}
meta.dist$frequencies |> ggplot(aes(x = nr_ipm, y = bin_unicode)) + geom_col()
```


```{r}
meta.dist <- str_interp("${url_api}/mmda/discourseme/${focus.discourseme.id}/description/${focus.discourseme.description.id}/meta") |> 
  request() |> 
  req_url_query(level = "protocol", key = "year", page_size = 1000) |> 
  req_auth_bearer_token(access.token) |> 
  req_perform() |> 
  magrittr::extract2("body") |>
  rawToChar() |> fromJSON()
```

```{r}
meta.dist$frequencies |> ggplot(aes(x = nr_ipm, y = bin_unicode)) + geom_col()
```


## constellation

- we create a constellation (project) **Befreiung** (and directly add focus discourseme)

```{r}
constellations <- str_interp("${url_api}/mmda/constellation/") |> 
  request() |> 
  req_auth_bearer_token(access.token) |> 
  req_perform() |> 
  magrittr::extract2("body") |>
  rawToChar() |> fromJSON()

constellations
```


```{r}
constellation <- str_interp("${url_api}/mmda/constellation/") |> 
  request() |> 
  req_body_json(list(discourseme_ids = list(focus.discourseme.id), name = "Befreiung")) |> 
  req_auth_bearer_token(access.token) |> 
  req_perform() |> 
  magrittr::extract2("body") |>
  rawToChar() |> fromJSON()

constellation.id <- constellation |> magrittr::extract2("id")
```

## collocation analysis

- how is **Befreiung** used in GermaParl?
- look at collocates
```{r}
constellation.description.id <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/") |> 
  request() |> 
  req_auth_bearer_token(access.token) |> 
  req_body_json(list(corpus_id = corpus.id, s = "s")) |> 
  req_perform() |> 
  magrittr::extract2("body") |> rawToChar() |> fromJSON() |> 
  magrittr::extract2("id")

collocation.id <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/collocation/") |> 
  request() |> 
  req_auth_bearer_token(access.token) |> 
  req_method("put") |> 
  req_body_json(list(focus_discourseme_id = focus.discourseme.id, p = "lemma", window = 10)) |> 
  req_perform() |> 
  magrittr::extract2("body") |> rawToChar() |> fromJSON() |> 
  magrittr::extract2("id")
```

### comparison of LRC and LLR

```{r}
coll.lrc <- request(str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/collocation/${collocation.id}/map")) |>
  req_auth_bearer_token(access.token) |>
  req_url_query(page_number = 1, page_size = 50, sort_by = "conservative_log_ratio", sort_order = "descending") |> 
  req_perform() |> 
  magrittr::extract2("body") |> rawToChar() |> fromJSON()

coll.llr <- request(str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/collocation/${collocation.id}/map")) |>
  req_auth_bearer_token(access.token) |>
  req_url_query(page_number = 1, page_size = 50, sort_by = "log_likelihood", sort_order = "descending") |> 
  req_perform() |> 
  magrittr::extract2("body") |> rawToChar() |> fromJSON()
```

- we display ranks here (score scales not comparable)
```{r, fig.width = 8, fig.height = 12}
overlap <- intersect(
  coll.llr |> magrittr::extract2("map") |>
    filter(source %in% c("items", "discourseme_unigram_items"), is.na(discourseme_id) | discourseme_id != focus.discourseme.id) |>
    arrange(desc(score)) |> head(50) |> pull("item"),
  coll.lrc |> magrittr::extract2("map") |> 
    filter(source %in% c("items", "discourseme_unigram_items"), is.na(discourseme_id) | discourseme_id != focus.discourseme.id) |> 
    arrange(desc(score)) |> head(50) |> pull("item")
)

rbind(
  coll.llr |> magrittr::extract2("map") |> 
    filter(source %in% c("items", "discourseme_unigram_items"), is.na(discourseme_id) | discourseme_id != focus.discourseme.id) |> arrange(desc(score)) |> head(50) |> 
    mutate(rank = 1:n(), am = "log likelihood"),
  coll.lrc |> magrittr::extract2("map") |>
    filter(source %in% c("items", "discourseme_unigram_items"), is.na(discourseme_id) | discourseme_id != focus.discourseme.id) |> arrange(desc(score)) |> head(50) |> 
    mutate(rank = 1:n(), am = "LRC")
) |> 
  select(item, am, rank, x, y) |> 
  mutate(overlap = item %in% overlap) |> 
  ggplot(aes(x = x, y = y, label = item, colour = overlap)) + 
  geom_label_repel(aes(size = - rank), max.overlaps = Inf, point.size = NA, min.segment.length = Inf) +
  xlab("") + ylab("") + 
  facet_wrap(~ am, ncol = 1) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
```

- LRC retrieves more content words
- additional words in LLR mostly highly frequent terms, including grammatical particles
- we use LRC here

### comparison of window sizes

- results for different window sizes (1:10)
```{r}
maps.windows <- tibble()
for (w in 1:10){

  collocation.id.w <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/collocation/") |> 
    request() |> 
    req_auth_bearer_token(access.token) |> 
    req_method("put") |> 
    req_body_json(list(focus_discourseme_id = focus.discourseme.id, p = "lemma", window = w, semantic_map_id = coll.lrc$semantic_map_id)) |> 
    req_perform() |> 
    magrittr::extract2("body") |> rawToChar() |> fromJSON() |> 
    magrittr::extract2("id")
  
  coll.lrc.w <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/collocation/${collocation.id.w}/map") |>
    request() |> 
    req_auth_bearer_token(access.token) |>
    req_url_query(page_number = 1, page_size = 50, sort_by = "conservative_log_ratio", sort_order = "descending") |> 
    req_perform() |> 
    magrittr::extract2("body") |> rawToChar() |> fromJSON()
  
  df <- coll.lrc.w$map
  df$window <- w
  
  maps.windows <- rbind(maps.windows, df)
}
```

- semantic maps for each window size
```{r, fig.height = 20, fig.width = 15}
maps.windows |> filter(source == "items", window %in% 1:6) |> 
  select(item, score, x, y, window) |> 
  ggplot(aes(x = x, y = y, label = item)) + 
  geom_label_repel(aes(size = score), max.overlaps = Inf, point.size = NA, min.segment.length = Inf) +
  xlab("") + ylab("") + 
  facet_wrap(~ window, ncol = 2) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
```

- low window sizes yield grammatical particles (articles, conjunctions, punctuation marks), attributions (adjectives), and transitive verbs
- high window sizes yield lemmata indicating subdiscourses or related discourses

- direct comparison between 3 and 10
```{r, fig.width = 8, fig.height = 12}
overlap <- intersect(
  maps.windows |> filter(source == "items", window == 3, score > 0) |> pull("item"),
  maps.windows |> filter(source == "items", window == 10, score > 0) |> pull("item")
)

rbind(
  maps.windows |> filter(window == 3),
  maps.windows |> filter(window == 10)
) |> 
  filter(source == "items", score > 0) |> 
  select(item, score, window, x, y) |> 
  mutate(overlap = item %in% overlap) |> 
  ggplot(aes(x = x, y = y, label = item, colour = overlap)) + 
  geom_label_repel(aes(size = score), max.overlaps = Inf, point.size = NA, min.segment.length = Inf) +
  xlab("") + ylab("") + 
  facet_wrap(~ window, ncol = 1) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
```

- we use a window size of 10 here, which captures related discourses

### comparison of parl. groups

- we compute "local" collocation profiles (i.e. compare co-occurrence frequencies with marginals in subcorpora)
- we use the same semantic map (based on global profile) for all analyses
```{r}
semantic_map.id <- coll.lrc$semantic_map_id
```

```{r}
factions <- c("CDU/CSU", "GRUENE", "SPD", "AfD")
sbcrpra <- factions

map.list <- tibble()
for (subcorpus in sbcrpra){

  print(subcorpus)

  subcorpus.id <- subcorpora |> filter(name == subcorpus) |> pull(id)

  if (length(subcorpus.id) > 0){

    const.description.subcorpus.id <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/") |>
      request() |> 
      req_auth_bearer_token(access.token) |>
      req_body_json(list(subcorpus_id = subcorpus.id, corpus_id = corpus.id, s = "s")) |>
      req_perform() |>
      magrittr::extract2("body") |> rawToChar() |> fromJSON() |>
      magrittr::extract2("id")

    # create collocation
    collocation.id.subcorpus <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${const.description.subcorpus.id}/collocation/") |>
      request() |> 
      req_method("put") |>
      req_auth_bearer_token(access.token) |>
      req_body_json(list(focus_discourseme_id = focus.discourseme.id, p = "lemma", window = 10, semantic_map_id = semantic_map.id, marginals = "local")) |>
      req_perform() |>
      magrittr::extract2("body") |> rawToChar() |> fromJSON() |>
      magrittr::extract2("id")

    # get items
    coll.subcorpus <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${const.description.subcorpus.id}/collocation/${collocation.id.subcorpus}/map") |>
      request() |> 
      req_auth_bearer_token(access.token) |>
      req_url_query(page_number = 1, page_size = 50, sort_by = "conservative_log_ratio", sort_order = "descending") |>
      req_perform() |>
      magrittr::extract2("body") |> rawToChar() |> fromJSON()

    map.subcorpus <- coll.subcorpus |> magrittr::extract2("map") |> mutate(subcorpus = subcorpus)

    map.list <- rbind(map.list, map.subcorpus)
  }
}
```

```{r}
maps.factions <- map.list |> filter(source == "items", score > 0)

maps.factions |>
  select(item, subcorpus, score) |> 
  pivot_wider(names_from = subcorpus, values_from = score) |> 
  mutate(sum.lrc = rowSums(across(where(is.numeric)), na.rm = T)) |>
  arrange(desc(sum.lrc))
```

```{r, fig.width = 10, fig.height = 15}
maps.factions |> 
  ggplot(aes(x = x, y = y, label = item)) + 
  geom_label_repel(aes(size = score), max.overlaps = Inf, point.size = NA, min.segment.length = Inf) +
  xlab("") + ylab("") + 
  facet_wrap(~ subcorpus, ncol = 1) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
```


- pairwise overlaps
```{r}
dfs <- tibble()
for (sbcrps1 in factions){
  for (sbcrps2 in factions){
    o <- overlap.pairwise(maps.factions |> filter(subcorpus == sbcrps1), maps.factions |> filter(subcorpus == sbcrps2))
    df <- tibble(left = sbcrps1, right = sbcrps2, overlap = o)
    dfs <- rbind(dfs, df)
  }
}

dfs |> 
  ggplot(aes(x = left, y = right)) + 
  geom_raster(aes(fill = overlap)) + 
  geom_text(aes(label = round(overlap, 2)), size = 2) +
  scale_fill_gradient(low = "grey90", high = "red")
```

## discourseme formation from collocates

- need to look at concordances to make sense of lemmata and find meaningful discoursemes
- TODO filtered concordancing is still ridiculously slow, somebody should really fix this …
```{r}
# conc <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/concordance/") |> 
#   request() |> 
#   # req_url_query(focus_discourseme_id = focus.discourseme.id, filter_discourseme_ids = c(2)) |> 
#   req_url_query(focus_discourseme_id = focus.discourseme.id, filter_item = "zusammen") |> 
#   req_auth_bearer_token(access.token) |> 
#   req_perform() |> 
#   magrittr::extract2("body") |>
#   rawToChar() |> fromJSON()
# 
# conc |> conc.kwic()
```

- this is a highly interactive procedure, cannot protocol each and every step
- in practice, categorise via drag & drop in interactive semantic maps of MMDA toolkit
- here, we can e.g. create one large list and categorise in LibreOffice or alike

```{r}
bind_rows(
  coll.lrc$map |> filter(source == "items") |> 
    mutate(source = "global", rank = 1:n()) |>
    filter(score > 0) |> 
    select(item, rank, score, source),
  maps.windows |> filter(source == "items") |>
    group_by(window) |> 
    filter(score > 0) |> 
    mutate(rank = 1:n()) |> 
    ungroup() |> 
    mutate(source = str_c("global_window-", window)) |>
    select(item, rank, score, source),
  map.subcorpus |> filter(source == "items") |>
    filter(score > 0) |> 
    group_by(subcorpus) |> 
    mutate(rank = 1:n()) |> 
    ungroup() |> 
    mutate(source = str_c("faction-", subcorpus)) |>
    select(item, rank, score, source)
) |> 
  pivot_wider(id_cols = item, names_from = source, values_from = c(rank, score), names_glue = "{.value}_{source}") |> 
  write_tsv("befreiung-collocates-raw.tsv")
```

- then read categorisation and do some visualisations / statistical analyses
```{r}
# discoursemes <- read_tsv("befreiung-collocates-cat.tsv") |> 
#   select(item, discourseme) |> 
#   drop_na()
```

```{r}
# discoursemes |> group_by(discourseme) |> summarise(n = n()) |> 
#   arrange(desc(n))
```

```{r}
# discoursemes |>
#   arrange(discourseme) 
```

- don't forget to feed back discoursemes into database
```{r}
# discoursemes |> 
#   mutate(name = discourseme) |> 
#   group_by(discourseme) |> 
#   group_walk(~ {
#     template_data <- .x |> 
#       mutate(p = "lemma", surface = item) |> 
#       select(p, surface) |> 
#       nest(template = c(p, surface)) |> 
#       pull(template)
#     
#     print(.x)
# 
#     body <- list(
#       name = unique(.x$name),
#       template = template_data[[1]]
#     )
# 
#     resp <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/discourseme-description") |>
#       request() |>
#       req_auth_bearer_token(access.token) |>
#       req_body_json(body) |>
#       req_perform() |>
#       magrittr::extract2("body") |> rawToChar() |> fromJSON()
#     print(resp)
#   })
```

- retrieve all discoursemes
```{r}
# discoursemes <- request(str_interp("${url_api}/mmda/discourseme/")) |>
#   req_auth_bearer_token(access.token) |>
#   req_perform() |> 
#   magrittr::extract2("body") |> rawToChar() |> fromJSON() |> 
#   tibble()
```

## constellation analysis

### global

- look at top-500 in global collocation analysis with discourseme categorisation
```{r}
# coll.const <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/collocation/${collocation.id}/map") |>
#   request() |> 
#   req_auth_bearer_token(access.token) |>
#   req_url_query(page_number = 1, page_size = 500, sort_by = "conservative_log_ratio", sort_order = "descending") |> 
#   req_perform() |> 
#   magrittr::extract2("body") |> rawToChar() |> fromJSON()
```

- visualisation of discoursemes: centering effect
```{r}
# coll.const |> 
#   magrittr::extract2("map") |> 
#   filter(source == "discoursemes") |> 
#   rename(conservative_log_ratio = score) |> 
#   left_join(discoursemes |> rename(discourseme_id = id, discourseme = name) |> select(discourseme_id, discourseme)) |> 
#   filter(discourseme_id != focus.discourseme.id, conservative_log_ratio > 0) |> 
#   ggplot(aes(x = x, y = y, label = item, col = discourseme)) + 
#   geom_label_repel(aes(size = conservative_log_ratio),
#                    max.overlaps = Inf, point.size = NA, min.segment.length = Inf, 
#                    max.time = 2, max.iter = 10e5) +
#   coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1)) +
#   xlab("") + ylab("") +
#   theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
#   guides(colour = "none")
```

- visualisation of items and discourseme items
```{r, fig.width = 35, fig.height = 15}
# coll.const |> 
#   magrittr::extract2("map") |> 
#   filter(source %in% c("items", "discourseme_items")) |> 
#   left_join(discoursemes |> rename(discourseme_id = id, discourseme = name) |> select(discourseme_id, discourseme)) |> 
#   rename(conservative_log_ratio = score) |> 
#   filter((is.na(discourseme_id) & conservative_log_ratio > 0) | (discourseme_id != focus.discourseme.id)) |> 
#   # write_tsv("climate-change-collocates.tsv") 
#   ggplot(aes(x = x, y = y, label = item, col = discourseme)) + 
#   geom_label_repel(aes(size = conservative_log_ratio),
#                    max.overlaps = Inf, point.size = NA, min.segment.length = Inf, 
#                    max.time = 2, max.iter = 10e5) +
#   coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1)) +
#   xlab("") + ylab("") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())
```


### subcorpora

- get all constellation descriptions
```{r}
# const.description.ids <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/") |>
#   request() |> 
#   req_auth_bearer_token(access.token) |>
#   req_perform() |> 
#   magrittr::extract2("body") |> rawToChar() |> fromJSON() |> tibble()
# 
# const.description.ids <- const.description.ids |>
#   left_join(subcorpora |> select(id, name) |> rename(subcorpus_id = id, subcorpus_name = name))
```

### LPs

- add discoursemes to constellation descriptions in LPs
```{r}
# discourseme.ids <- coll.const |> magrittr::extract2("map") |> drop_na() |> pull(discourseme_id) |> unique()
# sbcrpra <- str_c("LP", as.character(15:19))
# map.list <- tibble()
# for (subcorpus in sbcrpra){
# 
#   print(subcorpus)
# 
#   const.desc.id <- const.description.ids |> filter(subcorpus_name == subcorpus) |> tail(1) |> pull(id)
# 
#   # make sure discourseme descriptions are part of contellation description
#   str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${const.desc.id}/add-discoursemes") |>
#     request() |> 
#     req_method("patch") |>
#     req_auth_bearer_token(access.token) |>
#     req_body_json(list(discourseme_ids = discourseme.ids)) |>
#     req_perform() |>
#     magrittr::extract2("body") |> rawToChar() |> fromJSON() |>
#     magrittr::extract2("id")
# 
#   # get collocation
#   collocation.id.subcorpus <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${const.desc.id}/collocation/") |>
#     request() |> 
#     req_method("put") |>
#     req_auth_bearer_token(access.token) |>
#     req_body_json(list(focus_discourseme_id = focus.discourseme.id, p = "lemma", window = 10, semantic_map_id = semantic_map.id, marginals = "local")) |>
#     req_perform() |>
#     magrittr::extract2("body") |> rawToChar() |> fromJSON() |>
#     magrittr::extract2("id")
# 
#   # get items
#   coll.subcorpus <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${const.desc.id}/collocation/${collocation.id.subcorpus}/map") |>
#     request() |> 
#     req_auth_bearer_token(access.token) |>
#     req_url_query(page_number = 1, page_size = 100, sort_by = "conservative_log_ratio", sort_order = "descending") |>
#     req_perform() |>
#     magrittr::extract2("body") |> rawToChar() |> fromJSON()
#     map.subcorpus <- coll.subcorpus |> magrittr::extract2("map") |> mutate(subcorpus = subcorpus)
# 
#   map.list <- rbind(map.list, map.subcorpus)
# }
```

```{r, fig.width = 10, fig.height = 15}
# map.list |> 
#   filter(source %in% c("items", "discourseme_items")) |> 
#   left_join(discoursemes |> rename(discourseme_id = id, discourseme = name) |> select(discourseme_id, discourseme)) |> 
#   rename(conservative_log_ratio = score) |> 
#   filter(conservative_log_ratio > 0) |> 
#   filter((is.na(discourseme_id) & conservative_log_ratio > 0) | (discourseme_id != focus.discourseme.id)) |> 
#   ggplot(aes(x = x, y = y, label = item, col = discourseme)) + 
#   geom_label_repel(aes(size = conservative_log_ratio),
#                    max.overlaps = Inf, point.size = NA, min.segment.length = Inf, 
#                    max.time = 2, max.iter = 10e5) +
#   coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1)) +
#   xlab("") + ylab("") +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
#   facet_wrap(~ subcorpus, ncol = 1)
```

- semantic maps are ideal for extending discoursemes
```{r}
# conc <- str_interp("${url_api}/mmda/constellation/${constellation.id}/description/${constellation.description.id}/concordance/") |> 
#   request() |> 
#   req_url_query(focus_discourseme_id = focus.discourseme.id, filter_discourseme_ids = 3, filter_discourseme_ids = 4) |> 
#   req_auth_bearer_token(access.token) |> 
#   req_perform() |> 
#   magrittr::extract2("body") |>
#   rawToChar() |> fromJSON()
# 
# conc |> conc.kwic()
```
