---
title: "UFA"
editor: source
author: Philipp Heinrich
date: June 06, 2025
format:
  html:
    toc: true
    toc-float: true
    collapsed: false
    smooth-scroll: false
    code-fold: true
    number-sections: true
    df-print: paged
    embed-resources: True
---

## setup

- a couple of helper functions are defined in cwb-cads/R-utils
```{r, message = FALSE}
source("collocation.R")
source("concordance.R")
source("association.R")
source("api.R")
```

### API

- dev server
```{r}
.cads.password <- Sys.getenv("CWB_CADS_PASSWORD")
.cads.url <- "https://corpora.linguistik.uni-erlangen.de/cwb-cads-dev/"

cads_connect(user = "admin", password = .cads.password, url = .cads.url)
```

```{r}
Sys.time()
```

### corpus

```{r}
(corpora <- cads_list_corpora())
```

```{r}
cads_activate_corpus("TAGESSCHAU_2020_2022")
```

### meta data

```{r}
cads_corpus_metadata()
```


## subcorpus collection
```{r}
collection <- cads_create_subcorpus_collection("months", "text", "date", time_interval = "month")
```

## focus discourseme

### creation

- create focus discourseme **Wirtschaft**
```{r}
focus.discourseme <- cads_create_discourseme("Wirtschaft")
```

- and corresponding discourseme description (= realisation in corpus)
```{r}
focus.discourseme.description <- cads_create_discourseme_description(focus.discourseme$id, c("Wirtschaft"))
```

### breakdown

```{r}
cads_discourseme_breakdown(focus.discourseme$id, focus.discourseme.description$id)
```


### frequencies across months

```{r}
focus.meta <- cads_discourseme_meta(focus.discourseme$id, focus.discourseme.description$id, "text", "month")

focus.meta$frequencies |> tibble()
  mutate(month = lubridate::floor_date(as.Date(bin_datetime), "month")) |> 
  group_by(month) |> summarise(nr_matches = sum(nr_matches)) |> 
  ggplot(aes(y = nr_matches, x = month)) + geom_col()
```

## constellation

- create a constellation **Wirtschaft** (and directly add focus discourseme)

```{r}
constellation <- cads_create_constellation("Wirtschaft", list(focus.discourseme$id))
constellation.description <- cads_create_constellation_description(constellation$id, "s")
```


## usage fluctuation analysis

```{r}
collection.descriptions <- cads_create_constellation_description_collection(constellation$id, collection$id, "s")
```

```{r}
collection.ufa <- cads_constellation_ufa(constellation$id, collection.descriptions$id, focus.discourseme$id)


collection.ufa <- cads_constellation_ufa(6, 40, 27)
```



```{r}
collection.ufa$ufa |>
  unnest_wider(score_confidence) |> 
  mutate(x_label = as.Date(str_c(x_label, "-01"))) |> 
  ggplot(aes(x = x_label, y = score)) + geom_point() + xlab("") + 
  geom_line(aes(y = upper_95)) +
  geom_line(aes(y = lower_95)) +
  geom_line(aes(y = smooth)) +
  ggtitle("Usage Fluctuation Analysis") +
  stat_smooth(method = "gam", formula = y ~ s(x, bs = "cr", fx = FALSE, k = 5), linewidth = 1, fill="#707070", level = 0.95) +
  stat_smooth(method = "gam", formula = y ~ s(x, bs = "cr", fx = FALSE, k = 5), linewidth = 1, fill="#FFFF00", level = 0.99)
```


